---
title: "Testing"
output: html_notebook
---

# EXPLORE EUROSTAT LIBRARY 

Find data on eurostat package

```{r}
library(tidyverse)
library(rvest)
library(eurostat)


# Get Eurostat data listing
toc <- get_eurostat_toc()
```


List all folders

```{r}
# Check the first items
library(knitr)
kable(head(toc, n = 20))
```


Search and list folders

```{r}
# info about passengers
kable(head(search_eurostat("passenger transport")))
```


## _Download data - Modal split of passenger transport

```{r}
# For the original data, see
# http://ec.europa.eu/eurostat/tgm/table.do?tab=table&init=1&plugin=1&language=en&pcode=ts id <- search_eurostat("Modal split of passenger transport",
id <- search_eurostat("Modal split of passenger transport",
type = "table")$code[1]
dat <- get_eurostat(id, time_format = "num")
# Label the data
dat <- label_eurostat(dat)

```


## _Download data - Consumers indicators

```{r}
# For the original data, see
# http://ec.europa.eu/eurostat/tgm/table.do?tab=table&init=1&plugin=1&language=en&pcode=ts id <- search_eurostat("Modal split of passenger transport",

dat <- get_eurostat("ei_bsco_m", time_format = "num")
dat <- label_eurostat(dat)

```



## _Download data - GDP

```{r}
# For the original data, see
dat <- get_eurostat("namq_10_gdp", time_format = "num")
dat <- label_eurostat(dat)

```

Choose filters
- Current prices, million euro
- Unadjusted data (i.e. neither seasonally adjusted nor calendar adjusted data)
- Gross domestic product at market prices




